Statistical Reporting Inconsistencies in Experimental Linguistics

### 2025-05-15 manuscript/manuscript_anonymized.tex
minor fixes
Revisions and new analysis

\abstract{The present paper investigates the prevalence of statistical
reporting consistencies across articles in eight experimental
linguistics journals published between 2000 and 2023. Using the R
package ``Statcheck'', we retrieved 39532 statistical test and assessed
whether p-values were consistent with their test statistic and degrees
of freedom. Half of all articles that used null hypothesis signficance
testing contained at least one inconsistent p-value. One in eight
articles contained an inconsistency that may have affected the
statistical conclusion. The inconsistency rates were comparable across
journals and seem stable over publication years. We discuss possible
reasons for this high rate and offer actionable steps for authors,
reviewers, and editors to remedy this state-of-affairs.}

@@

What we know about human language and its cognitive underpinnings is
often informed by experimental data. Researchers test theoretical
predictions with their data using statistical tests. Depending on the
results of their tests, researchers make claims for or against
theoretical assumptions.\\
Since these tests play such an integral part in the argumentation
process of experimentalists, both the data these tests are based on and
the computational procedure of the tests itself should be transparent in
order for other researchers to critically evaluate them. But more than
that, things can go wrong. Humans are fallible; they make mistakes.
Transparent sharing allows others to detect and correct human error. In
the recent years, quantitative linguistics have seen repeated calls to
become more transparent and reproducible by sharing data and statistical
protocols (\citeproc{ref-arvan2022reproducibility}{Arvan et al., 2022};
\citeproc{ref-laurinavichyute2022share}{Laurinavichyute et al., 2022};
\citeproc{ref-roettger2019researcher}{Roettger, 2019}). Despite these
calls, sharing of statistical protocols is still rather rare across the
language sciences (\citeproc{ref-bochynska2023reproducible}{Bochynska et
al., 2023}). If statistical procedures cannot be critically evaluated,
human errors might be left undetected and thus remain uncorrected in the
publication record. And if undetected errors affect the decision
procedure of the analysis, i.e.~whether a hypothesis is rejected or
accepted, these errors might lead to -at best- overconfident, -at worst-
false theoretical conclusions. The present paper will present evidence
that the published literature in experimental linguistics contains a
concerning amount of such statistical errors, a state of affairs which
warrants more rigorous data sharing practices.

@@

The consistency of these values can be automatically assessed if
statistical tests are reported in an unambiguous format. Recently, a
series of studies used such automatic assessments to evaluate the
prevalence of inconsistent statistical reporting in psychology
(\citeproc{ref-bakker2011mis}{Bakker \& Wicherts, 2011},
\citeproc{ref-bakker2014outlier}{2014};
\citeproc{ref-caperos2013consistency}{Caperos \& Pardo, 2013};
\citeproc{ref-claesen2023data}{Claesen et al., 2023};
\citeproc{ref-green2018statcheck}{Green et al., 2018};
\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016};
\citeproc{ref-nuijten2020statcheck}{Nuijten \& Polanin, 2020};
\citeproc{ref-veldkamp2014statistical}{Veldkamp et al., 2014};
\citeproc{ref-wicherts2011willingness}{Wicherts et al., 2011}), medical
sciences (\citeproc{ref-garcia2004incongruence}{Garcı́a-Berthou \&
Alcaraz, 2004}; \citeproc{ref-van2023comparing}{Van Aert et al., 2023}),
psychiatry (\citeproc{ref-berle2007inconsistencies}{Berle \& Starcevic,
2007}), cyber security studies (\citeproc{ref-gross2021fidelity}{Groß,
2021}), technological education research
(\citeproc{ref-buckley2023estimating}{Buckley et al., 2023}), and
experimental philosophy (\citeproc{ref-colombo2018statistical}{Colombo
et al., 2018}). For example, looking at over 250 thousand p-values
published in major psychology journals, Nuijten et al.
(\citeproc{ref-nuijten2016prevalence}{2016}) found that around 50\% of
the articles with statistical results contained at least one
inconsistency and around 13\% contained at least one decision
inconsistency. Other studies report on inconsistency rates between 4\%
and 14\%, with between 10\% and 63\% of articles containing at least one
inconsistency and between 3\% and 21\% contain at least one decision
consistency. These high inconsistency rates in other disciplines are
concerning and have led to a constructive dialog, resulting in
recommendations for best practice to either avoid inconsistencies or
make them more detectable.

To assess the prevalence of statistical-reporting inconsistencies in
experimental linguistics, the present paper conceptually replicates
Nuijten et al. (\citeproc{ref-nuijten2016prevalence}{2016}) and assesses
p-values reported in eight experimental linguistic journals published
between 2000 and 2023. We explore whether the rate of inconsistency
differs across journals, whether their rate has changed over the course
of the last 23 years and whether there is evidence for bias in these
statistical-reporting inconsistencies. We discuss the results and offer
concrete recommendations for authors, reviewers, and editors to tackle
this problem.

@@

All quantitative analyses were conducted using R Core Team
(\citeproc{ref-rmanual}{2021}). All derived data and corresponding R
scripts are publicly available here:
\url{https://github.com/timo-b-roettger/ErrorsLing}

\subsection{3.2. Statcheck}\label{statcheck}

We used the R package Statcheck (Version 1.4.1-beta.2,
\citeproc{ref-nuijten2023statcheck}{Nuijten \& Epskamp, 2023}) to
automatically detect statistical-reporting inconsistencies. Statcheck
works as follows: After converting articles in pdf or html format to
plain text, Statcheck searches for specific strings that correspond to a
NHST result using regular expressions. That way, Statcheck can detect
results of t-tests, F-tests, Z-tests, χ2-tests, correlation tests, and Q
tests as long as the test result fulfills three conditions: (a) the test
result is reported completely including the test statistic, the degrees
of freedom (if applicable), and the p-value; (b) the test result is in
the body of the text, i.e. Statcheck usually misses information in
tables; and (c) the test result is reported in American Psychological
Association style (\citeproc{ref-APA2020}{American Psychological
Association, 2020}). Given these constraints, Statcheck is estimated to
detect roughly 60\% of all reported NHST results
(\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). Statcheck
uses the reported test statistic and degrees of freedom to recalculate
the p-value, compares the reported and recalculated p-value and, if
there is a mismatch, flags the test as containing an ``inconsistency.''
The algorithm takes into account that tests might have been performed as
one-tailed by identifying the search strings ``one-tailed,''
``one-sided,'' or ``directional'' in the body of the text. Moreover,
Statcheck considers p = .000 and p \textless{} .000 as inconsistent
because p-values of exactly zero are mathematically impossible and the
APA manual (\citeproc{ref-APA2020}{American Psychological Association,
2020}) advises to report very small p-values as p \textless{} .001.
Validity checks of Statcheck suggest that inter-rater reliability
between manual coding and Statcheck is high, i.e.~0.76 for
inconsistencies and 0.89 for decision inconsistencies
(\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). The
overall accuracy of Statcheck is estimated to be between 96.2\% to
99.9\% (\citeproc{ref-nuijten2017validity}{Nuijten et al., 2017}). We
thus consider Statcheck a valid proxy of the prevalence of statistical
reporting inconsistencies.

\subsection{3.3. Sample}\label{sample}

Focusing on experimental linguistic research, we used Kobrock and
Roettger (\citeproc{ref-kobrock2023assessing}{2023}) as a point of
departure who list 100 linguistic journals that had at least a hundred
articles published at the time of assessment (2021) and a high ratio of
articles containing the search string ``experiment*'' in title, abstract
and/or keywords. Out of these 100 journals, we selected all journals
with at least 10\% of articles containing the search string
``experiment*''. Out of the remaining 37 journals, we selected those
journals that urged APA formatting either in the main body of the text
or specifically regarding statistics in the author guidelines, resulting
in nine remaining journals. Moreover, to access the articles in .pdf
format, the articles had to be either accessible to us through our
library license, or open access, resulting in a final list of eight
journals: Applied Psycholinguistics (APS), Bilingualism: Language and
Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language
and Speech (LaS), Language Learning and Techology (LLT), Journal of
Language and Social Psychology (LSP), Journal of Child Language (JCL),
and Studies in Second Language Acquisition (SLA).

We included only original research articles within the publication years
of 2000-2023, excluding any book reviews, response articles,
commentaries, editorials, corrigenda, errata, advertisements, etc.
Articles from LAB spanned 2011-2023, while the other journals spanned
2000-2023. Statcheck could not parse 157 articles, likely related to
issues with rendering the Chi-Squared symbol being erroneously converted
from .pdf to .txt. No assessable statistics were found by Statcheck in
JCL articles from 2000 to 2002 and in SLA articles from 2004 to 2008.
This procedure resulted in 5804 research articles which were submitted
to analysis.

@@

\subsection{4.1 Prevalence of
inconsistencies}\label{prevalence-of-inconsistencies}

The results are summarized in Table~\ref{tbl-summary}. Out of 5804
articles, 3059 articles contained statistical tests that statcheck could
assess (53\%), amounting to 39532 assessable p-values. 5166 p-values
were flagged as inconsistent (13.1\%) out of which 655 were considered
decision inconsistencies (1.7\%) (see Table~\ref{tbl-summary}).

@@

{\caption{{Number of eligible articles, assessable articles and results,
inconsistencies and decision inconsistencies across all journals.
(Applied Psycholinguistics (APS), Bilingualism: Language and Cognition
(BLC), Linguistic Approaches to Bilingualism (LAB), Language and Speech
(LaS), Language Learning and Techology (LLT), Journal of Language and
Social Psychology (LSP), Journal of Child Language (JCL), and Studies in
Second Language Acquisition (SLA))}{\label{tbl-summary}}}}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.0755}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1698}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1887}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1792}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1509}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2358}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Journal
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
eligible articles
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
assessable articles
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
assessable results
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
inconsistencies
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
decision inconsistencies
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
APS & 953 & 690 & 9570 & 1368 & 170 \\
BLC & 964 & 610 & 9093 & 1161 & 120 \\
JCL & 1109 & 529 & 6240 & 750 & 69 \\
LAB & 471 & 133 & 1719 & 234 & 25 \\
LLT & 421 & 111 & 919 & 201 & 61 \\
LSP & 695 & 376 & 4320 & 429 & 60 \\
LaS & 598 & 363 & 4717 & 552 & 86 \\
SLA & 593 & 247 & 2954 & 471 & 64 \\
Total & 5804 & 3059 & 39532 & 5166 & 655 \\
\end{longtable}

@@

The proportion of inconsistencies ranged from 10 to 22\% across journals
(1 to 7\% for decision inconsistencies) (see
Figure~\ref{fig-stacked-bar}). These rates appear to be stable across
year of publication (see Figure~\ref{fig-year-prop}). On average, 50\%
of assessable articles contained one or more inconsistencies (journals
range from 43 to 56\%) and 13\% contained one or more decision
inconsistencies (journals range from 11 to 15\%).

@@

\caption{\label{fig-stacked-bar}Proportion of articles containing at
least one inconsistency / decision inconsistency. (Applied
Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC),
Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS),
Language Learning and Techology (LLT), Journal of Language and Social
Psychology (LSP), Journal of Child Language (JCL), and Studies in Second
Language Acquisition (SLA))}

@@

\caption{\label{fig-year-prop}Proportion of inconsistencies / decision
inconsistencies plotted from 2000 to 2023. Left: rates pooled across
journals; Right: rates per journal. (Applied Psycholinguistics (APS),
Bilingualism: Language and Cognition (BLC), Linguistic Approaches to
Bilingualism (LAB), Language and Speech (LaS), Language Learning and
Techology (LLT), Journal of Language and Social Psychology (LSP),
Journal of Child Language (JCL), and Studies in Second Language
Acquisition (SLA))}

@@

When examining reported against recalculated p-values for all
inconsistencies (see Figure~\ref{fig-scatter}), we can identify certain
spatial patterns (only inconsistencies that are based on `=' comparisons
were included to make them easier to interpret, n = 3779 out of 5166).
First, the center of density of points is located in the bottom left
corner with more p-values reported closer to the most common alpha level
of 0.05. This is not surprising, since the quantitative sciences exhibit
a documented publication bias, with hypotheses being more often
confirmed (and thus p \textless{} .05) than not
(\citeproc{ref-franco2014publication}{Franco et al., 2014};
\citeproc{ref-sterling1959publication}{Sterling, 1959}). Second, there
is a linear density of points along the diagonal axis which corresponds
to numerically small inconsistencies, some of which might be related to
simple rounding errors. However, comparing the diagonal to the black
line, which represents a linear model predicting recalculated by
reported p-values, we can see a clear divergence of what is expected if
inconsistencies were equally likely in both directions. The slope of the
regression line is flatter than the diagonal axis which means that, on
average, inconsistencies have a tendency to exhibit smaller reported
p-values than their recalculated counterparts. Similarly,
inconsistencies that report the p-value as being larger or smaller than
a reference value (e.g.~p \textgreater{} 0.05 or p \textless{} 0.05 )
are not equally prevalent. There were 4.4\% of inconsistencies with p
being reported as larger than a reference but 7.4\% inconsistencies with
p being reported as smaller than a reference (e.g.~p \textless{} 0.05).
So even if we assumed these inconsistencies were merely typos of the
comparison sign, inconsistencies that erroneously report the p-value to
be smaller than a reference value are more frequent than inconsistencies
that erroneously report the p-value to be larger than a reference value.
These biases are also reflected in decision inconsistencies. Of all
decision inconsistencies (n = 655), 71\% represent cases in which a
reported significant result (p \textless{} 0.05) is recalculated as
non-significant (p \textgreater{} 0.05).

@@

\caption{\label{fig-scatter}Reported vs.~recalculated p-values for
inconsistencies that are based on `=' comparions. Black line
descriptively indicates the linear relationship, indicating a bias
towards lower reported p-values.}

@@

The present study found a large amount of statistical reporting
inconsistencies across a sample of 5804 experimental linguistic
articles, containing 39532 assessable p-values. 13.1\% of all p-values
were flagged as inconsistent and 1.7\% were flagged as decision
inconsistencies, i.e.~the reported p-value is on the opposite side of
the alpha threshold than the recalculated p-value. On average, 50\% of
assessable articles contained at least one inconsistency and 13\%
contained at least one decision inconsistency.

The present study can be considered a conceptual replication of previous
studies investigating statistical-reporting inconsistency across
different disciplines (\citeproc{ref-bakker2011mis}{Bakker \& Wicherts,
2011}, \citeproc{ref-bakker2014outlier}{2014};
\citeproc{ref-caperos2013consistency}{Caperos \& Pardo, 2013};
\citeproc{ref-veldkamp2014statistical}{Veldkamp et al., 2014};
\citeproc{ref-wicherts2011willingness}{Wicherts et al., 2011}) and most
recent such assessment using Statcheck
(\citeproc{ref-buckley2023estimating}{Buckley et al., 2023};
\citeproc{ref-colombo2018statistical}{Colombo et al., 2018};
\citeproc{ref-gross2021fidelity}{Groß, 2021};
\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). The
discovered rates fall in line with these studies who report on
inconsistency rates between 4\% and 14\%, with between 10\% and 63\% of
articles containing at least one inconsistency.

@@

@@

@@

>>>>

\abstract{The present paper investigates the prevalence of statistical
reporting consistencies across articles in eight experimental
linguistics journals published between 2000 and 2023. Using the R
package ``Statcheck'', we retrieved 85,442 statistical test from 13,065
articles and assessed whether p-values were consistent with their test
statistic and degrees of freedom. Half of all articles (50\%) that used
null hypothesis significance testing contained at least one inconsistent
p-value. More than one in eight articles (13\%) contained an
inconsistency that may have affected the statistical conclusion. The
inconsistency rates were comparable across journals and seem stable over
publication years. We discuss possible reasons for this high rate and
offer actionable steps for authors, reviewers, and editors to remedy
this state-of-affairs.}

@@

What we know about human language and its cognitive underpinnings is
often informed by experimental data. Researchers test theoretical
predictions with their data using statistical tests. Depending on the
results of their tests, researchers make claims for or against
theoretical assumptions.\\
Since these tests play such an integral part in the argumentation
process of experimentalists, both the data these tests are based on and
the computational procedure of the tests itself should be transparent in
order for other researchers to critically evaluate them. But more than
that, things can go wrong. Humans are fallible; they make mistakes.
Transparent sharing allows others to detect and correct human error. In
recent years, quantitative linguistics have seen repeated calls to
become more transparent and reproducible by sharing data and statistical
protocols, often under the banner of ``open science''
(\citeproc{ref-arvan2022reproducibility}{Arvan et al., 2022};
\citeproc{ref-laurinavichyute2022share}{Laurinavichyute et al., 2022};
\citeproc{ref-roettger2019researcher}{Roettger, 2019}). Despite these
calls, sharing of statistical protocols is still rather rare across the
language sciences (\citeproc{ref-bochynska2023reproducible}{Bochynska et
al., 2023}). If statistical procedures cannot be critically evaluated,
human errors might be left undetected and thus remain uncorrected in the
publication record. And if undetected errors affect the decision
procedure of the analysis, i.e.~whether a hypothesis is rejected or
accepted, these errors might lead to -at best- overconfident, -at worst-
false theoretical conclusions. The present paper will present evidence
that the published literature in experimental linguistics contains a
concerning amount of such statistical errors, a state of affairs which
warrants more rigorous data sharing practices.

@@

The consistency of these values can be automatically assessed if
statistical tests are reported in an unambiguous format. Recently, a
series of studies used such automatic assessments to evaluate the
prevalence of inconsistent statistical reporting in psychology
(\citeproc{ref-bakker2011mis}{Bakker \& Wicherts, 2011},
\citeproc{ref-bakker2014outlier}{2014};
\citeproc{ref-caperos2013consistency}{Caperos \& Pardo, 2013};
\citeproc{ref-claesen2023data}{Claesen et al., 2023};
\citeproc{ref-green2018statcheck}{Green et al., 2018};
\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016};
\citeproc{ref-nuijten2020statcheck}{Nuijten \& Polanin, 2020};
\citeproc{ref-veldkamp2014statistical}{Veldkamp et al., 2014};
\citeproc{ref-wicherts2011willingness}{Wicherts et al., 2011}), medical
sciences (\citeproc{ref-garcia2004incongruence}{Garcı́a-Berthou \&
Alcaraz, 2004}; \citeproc{ref-van2023comparing}{Van Aert et al., 2023}),
psychiatry (\citeproc{ref-berle2007inconsistencies}{Berle \& Starcevic,
2007}), cyber security studies (\citeproc{ref-gross2021fidelity}{Groß,
2021}), technological education research
(\citeproc{ref-buckley2023estimating}{Buckley et al., 2023}), and
experimental philosophy (\citeproc{ref-colombo2018statistical}{Colombo
et al., 2018}). For example, looking at over 250,000 p-values published
in major psychology journals, Nuijten et al.
(\citeproc{ref-nuijten2016prevalence}{2016}) found that around 50\% of
the articles with statistical results contained at least one
inconsistency and around 13\% contained at least one decision
inconsistency. Other studies report on inconsistency rates between 4\%
and 14\%, with between 10\% and 63\% of articles containing at least one
inconsistency and between 3\% and 21\% contain at least one decision
consistency. These high inconsistency rates in other disciplines are
concerning and have led to a constructive dialog, resulting in
recommendations for best practice to either avoid inconsistencies or
make them more detectable.

To assess the prevalence of statistical-reporting inconsistencies in
experimental linguistics, the present paper conceptually replicates
Nuijten et al. (\citeproc{ref-nuijten2016prevalence}{2016}) and assesses
p-values reported in thirteen experimental linguistic journals published
between 2000 and 2023. We explore whether the rate of inconsistency
differs across journals, whether their rate has changed over the course
of the last 23 years and whether there is evidence for bias in these
statistical-reporting inconsistencies. We discuss the results and offer
concrete recommendations for authors, reviewers, and editors to tackle
this problem.

@@

All quantitative analyses were conducted using R Core Team
(\citeproc{ref-rmanual}{2025}). All derived data and corresponding R
scripts are available
\href{https://osf.io/gx3ub/?view_only=25957e06b5f449d6820b529eb0ae1937}{here}.

\subsection{3.2. Sample}\label{sample}

Focusing on experimental linguistic research, we used Kobrock and
Roettger (\citeproc{ref-kobrock2023assessing}{2023}) as a point of
departure who list 100 linguistic journals that had at least a hundred
articles published at the time of assessment (2021) and a high ratio of
articles containing the search string ``experiment*'' in title, abstract
and/or keywords. Out of these 100 journals, we selected all journals
with at least 10\% of articles containing the search string
``experiment*''. Out of the remaining 37 journals, we selected those
journals that urged APA formatting either in the main body of the text
or specifically regarding statistics in the author guidelines, resulting
in nine remaining journals. Moreover, to access the articles in .pdf
format, the articles had to be either accessible to us through our
library license, or open access, resulting in a final list of eight
journals: Applied Psycholinguistics (APS), Bilingualism: Language and
Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language
and Speech (LaS), Language Learning and Techology (LLT), Journal of
Language and Social Psychology (LSP), Journal of Child Language (JCL),
and Studies in Second Language Acquisition (SLA). An anonymous reviewer
raised the justified concern that the resulting sample might not
represent the vast majority of work in experimental linguistics. To
address this concern, we additionally included those five journals that
have published the highest absolute number of experimental articles,
regardless of whether they explicitly urge APA formatting or not,
resulting in the inclusion of the following psycholinguistic outlets:
Journal of Memory and Language (JML), Language Cognition and
Neuroscience (LCN, former Language and Cognitive Processes), Journal of
Psycholinguistic Research (JPR), Journal Of Speech Language And Hearing
Research (SLH), and Brain and Language (BAL).

We included only original research articles within the publication years
of 2000-2023, excluding any book reviews, response articles,
commentaries, editorials, corrigenda, errata, advertisements, etc.

\subsection{3.3. Statcheck}\label{statcheck}

We used the R package Statcheck (Version 1.5.0,
\citeproc{ref-nuijten2023statcheck}{Nuijten \& Epskamp, 2024}) to
automatically detect statistical-reporting inconsistencies. Statcheck
works as follows: After converting articles in pdf or html format to
plain text, Statcheck searches for specific strings that correspond to a
NHST result using regular expressions. That way, Statcheck can detect
results of t-tests, F-tests, Z-tests, χ2-tests, correlation tests, and Q
tests as long as the test result fulfills three conditions: (a) the test
result is reported completely including the test statistic, the degrees
of freedom (if applicable), and the p-value; (b) the test result is in
the body of the text, i.e. Statcheck usually misses information in
tables; and (c) the test result is reported in American Psychological
Association style (\citeproc{ref-APA2020}{American Psychological
Association, 2020}). Given these constraints, Statcheck is estimated to
detect roughly 60\% of all reported NHST results
(\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). Statcheck
uses the reported test statistic and degrees of freedom to recalculate
the p-value, compares the reported and recalculated p-value and, if
there is a mismatch, flags the test as containing an ``inconsistency.''
The algorithm takes into account that tests might have been performed as
one-tailed by identifying the search strings ``one-tailed,''
``one-sided,'' or ``directional'' in the body of the text. Moreover,
Statcheck considers p = .000 and p \textless{} .000 as inconsistent
because p-values of exactly zero are mathematically impossible and the
APA manual (\citeproc{ref-APA2020}{American Psychological Association,
2020}) advises to report very small p-values as p \textless{} .001.
Validity checks of Statcheck suggest that inter-rater reliability
between manual coding and Statcheck is high, i.e.~0.76 for
inconsistencies and 0.89 for decision inconsistencies
(\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). The
overall accuracy of Statcheck is estimated to be between 96.2\% to
99.9\% (\citeproc{ref-nuijten2017validity}{Nuijten et al., 2017}). We
thus consider Statcheck a valid proxy of the prevalence of statistical
reporting inconsistencies.

Articles from LAB spanned 2011-2023; LCN spanned 2015-2023. Statcheck
could not parse 291 articles, likely related to issues with rendering
the Chi-Squared symbol being erroneously converted from .pdf to .txt,
resulting in some gaps in the coverage. This procedure resulted in 13065
research articles which were submitted to analysis.

\subsection{3.4. Analysis and research
questions}\label{analysis-and-research-questions}

The aims of this study were explicitly exploratory and hypothesis
generating, thus analyses remain merely descriptive by reporting on the
proportion of articles/tests that are statistically (in)consistent. Our
investigation was set out to explore the following research questions:
(1) How prevalent are statistical inconsistencies (and decision
inconsistencies) in our sample? (2) Do inconsistency rates vary across
journals and/or their publication year? And (3) is there evidence for
bias, i.e.~are processes that result in inconsistencies more likely to
produce lower p-values?

@@

\subsection{4.1 Inconsistencies are highly
prevalent}\label{inconsistencies-are-highly-prevalent}

The results are summarized in Table~\ref{tbl-summary}. Out of 13065
articles, 6484 articles contained statistical tests that Statcheck could
assess (50\%), amounting to 85442 assessable p-values. Interestingly,
the five journals that did not explicitly encourage APA formatting had
virtually identical rates of assessable articles compared to the
original sample. Overall, 11279 p-values were flagged as inconsistent
(13.2\%) out of which 1442 were considered decision inconsistencies
(1.7\%) (see Table~\ref{tbl-summary}).

@@

{\caption{{Number of eligible articles, assessable articles and results,
inconsistencies and decision inconsistencies across all journals.
(Applied Psycholinguistics (APS), Language and Brain (BAL),
Bilingualism: Language and Cognition (BLC), Journal of Memory and
Language (JML), Journal of Psycholinguistic Research (JPR), Linguistic
Approaches to Bilingualism (LAB), Language and Speech (LaS), Language
Cognition and Neuroscience (LCN, former Language and Cognitive
Processes), Language Learning and Techology (LLT), Journal of Language
and Social Psychology (LSP), Journal of Child Language (JCL), and
Studies in Second Language Acquisition (SLA), Journal Of Speech Language
And Hearing Research (SLH))}{\label{tbl-summary}}}}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.0755}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1698}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1887}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1792}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1509}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2358}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Journal
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
eligible articles
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
assessable articles
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
assessable results
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
inconsistencies
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
decision inconsistencies
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
APS & 953 & 690 & 9570 & 1447 & 198 \\
BAL & 2253 & 780 & 10344 & 1524 & 198 \\
BLC & 964 & 610 & 9093 & 1223 & 134 \\
JCL & 1109 & 529 & 6240 & 779 & 76 \\
JML & 1507 & 768 & 13236 & 1004 & 132 \\
JPR & 1137 & 534 & 5576 & 848 & 97 \\
LAB & 471 & 133 & 1719 & 239 & 25 \\
LCN & 751 & 397 & 5979 & 909 & 95 \\
LLT & 421 & 111 & 919 & 210 & 64 \\
LSP & 695 & 376 & 4320 & 449 & 68 \\
LaS & 597 & 362 & 4708 & 564 & 88 \\
SLA & 593 & 247 & 2954 & 476 & 66 \\
SLH & 1614 & 947 & 10784 & 1607 & 201 \\
Total & 13065 & 6484 & 85442 & 11279 & 1442 \\
\end{longtable}

@@

\subsection{4.2 Inconsistencies are stable across journals and
publication
year}\label{inconsistencies-are-stable-across-journals-and-publication-year}

The proportion of inconsistencies ranged from 8 to 23\% across journals
(1 to 7\% for decision inconsistencies) (see
Figure~\ref{fig-stacked-bar}). These rates appear to be stable across
year of publication (see Figure~\ref{fig-year-prop}). On average, 50\%
of assessable articles contained one or more inconsistencies (journals
range from 45 to 56\%) and 13\% contained one or more decision
inconsistencies (journals range from 11 to 17\%).

@@

\caption{\label{fig-stacked-bar}Proportion of articles containing at
least one inconsistency / decision inconsistency. (Applied
Psycholinguistics (APS), Language and Brain (BAL), Bilingualism:
Language and Cognition (BLC), Journal of Memory and Language (JML),
Journal of Psycholinguistic Research (JPR), Linguistic Approaches to
Bilingualism (LAB), Language and Speech (LaS), Language Cognition and
Neuroscience (LCN, former Language and Cognitive Processes), Language
Learning and Techology (LLT), Journal of Language and Social Psychology
(LSP), Journal of Child Language (JCL), and Studies in Second Language
Acquisition (SLA), Journal Of Speech Language And Hearing Research
(SLH))}

@@

\caption{\label{fig-year-prop}Proportion of inconsistencies / decision
inconsistencies across time overall (left panel) and split into journals
(right panel). (Applied Psycholinguistics (APS), Language and Brain
(BAL), Bilingualism: Language and Cognition (BLC), Journal of Memory and
Language (JML), Journal of Psycholinguistic Research (JPR), Linguistic
Approaches to Bilingualism (LAB), Language and Speech (LaS), Language
Cognition and Neuroscience (LCN, former Language and Cognitive
Processes), Language Learning and Techology (LLT), Journal of Language
and Social Psychology (LSP), Journal of Child Language (JCL), and
Studies in Second Language Acquisition (SLA), Journal Of Speech Language
And Hearing Research (SLH))}

@@

\subsection{4.3 Inconsistencies are
biased}\label{inconsistencies-are-biased}

Inconsistencies that report the p-value as being larger or smaller than
a reference value (e.g.~p \textgreater{} 0.05 or p \textless{} 0.05,
respectively) are not equally prevalent in the sample: There were 4.4\%
of inconsistencies with p being reported as larger than a reference but
7.1\% inconsistencies with p being reported as smaller than a reference.
So even if we assumed these inconsistencies were merely typos of the
comparison sign (e.g.~\textless{} instead of \textgreater),
inconsistencies that erroneously report the p-value to be smaller than a
reference value are more frequent than inconsistencies that erroneously
report the p-value to be larger than a reference value.

These biases are also reflected in decision inconsistencies. Of all
decision inconsistencies (n = 1442), 72\% represent cases in which a
reported significant result (p \textless{} 0.05) is recalculated as
non-significant (p \textgreater{} 0.05), i.e.~non-significant results
are more than twice as likely to be erroneously reported as significant
than the other way around. The latter pattern, however, seems to be a
bias of the past. Reproducing Nuijten et al.
(\citeproc{ref-nuijten2016prevalence}{2016}), Figure~\ref{fig-scatter}
plots the development of the bias observed for decision inconsistencies
over time. The prevalence of decision inconsistencies in significant
p-values seems to have slightly decreased over the years, while the
prevalence of decision inconsistencies in non-significant p-values seems
to have slightly increased over the years.

@@

\caption{\label{fig-scatter}Percentage of decision inconsistencies
falsely reporting significance (black) or non-significance (orange),
plotted across publication years.}

@@

The present study found a large amount of statistical reporting
inconsistencies across a sample of 13065 experimental linguistic
articles, containing 85442 assessable p-values. 13.2\% of all p-values
were flagged as inconsistent and 1.7\% were flagged as decision
inconsistencies, i.e.~the reported p-value is on the opposite side of
the alpha threshold than the recalculated p-value. On average, 50\% of
assessable articles contained at least one inconsistency and 13\%
contained at least one decision inconsistency.

The present study can be considered a conceptual replication of previous
studies investigating statistical-reporting inconsistency across
different disciplines (\citeproc{ref-bakker2011mis}{Bakker \& Wicherts,
2011}, \citeproc{ref-bakker2014outlier}{2014};
\citeproc{ref-caperos2013consistency}{Caperos \& Pardo, 2013};
\citeproc{ref-veldkamp2014statistical}{Veldkamp et al., 2014};
\citeproc{ref-wicherts2011willingness}{Wicherts et al., 2011}) and most
recent assessments using Statcheck
(\citeproc{ref-buckley2023estimating}{Buckley et al., 2023};
\citeproc{ref-colombo2018statistical}{Colombo et al., 2018};
\citeproc{ref-gross2021fidelity}{Groß, 2021};
\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). The
discovered inconsistency rates fall in line with these studies that
report on rates between 4\% and 14\%, with between 10\% and 63\% of
articles containing at least one inconsistency.

@@

\bibitem[\citeproctext]{ref-cumming2014new}
Cumming, G. (2014). The new statistics: Why and how. \emph{Psychological
Science}, \emph{25}(1), 7--29.

\bibitem[\citeproctext]{ref-de2015cognitive}
De Bruin, A., Treccani, B., \& Della Sala, S. (2015). Cognitive
advantage in bilingualism: An example of publication bias?
\emph{Psychological Science}, \emph{26}(1), 99--107.

@@

\bibitem[\citeproctext]{ref-isbilen2022statistical}
Isbilen, E. S., \& Christiansen, M. H. (2022). Statistical learning of
language: A meta-analysis into 25 years of research. \emph{Cognitive
Science}, \emph{46}(9), e13198.

@@

\bibitem[\citeproctext]{ref-lehtonen2018bilingualism}
Lehtonen, M., Soveri, A., Laine, A., Järvenpää, J., De Bruin, A., \&
Antfolk, J. (2018). Is bilingualism associated with enhanced executive
functioning in adults? A meta-analytic review. \emph{Psychological
Bulletin}, \emph{144}(4), 394.
